Uppangala tree functional traits: Article draft

### 2016-07-22 inst/doc/draft2.Rmd
v0.1.9021 Pre tree data changes

Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...

@@

@@

* Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition. For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
* Leaf investment definition
* Leaf ressource acuisition definition
* Analyses scheme presentation:
    + Are functionnal composition of communities linked to the environment ?
        - Do covariations in functionnal traits illustrate ecological strategy axes ?
        - Are functionnal traits phylogenetically conserved ?
    + Are communities spatially structured ?
        - Are functionnal composition of communities spatially structured ?
        - Are phylogenetic composition of communities spatially structured ?

@@

The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996. The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998]. Each tree girth is monitored by a two-year census.
**Girth monitoring to develop**

@@

We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates. Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals. Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot. For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013. Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy. Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling. Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours. For some individuals we were unable to collect an appropriate branch sample (58 over 267 individuals representing 12 species over 91).

Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite]. Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time. Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias. The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues. Branch wood samples were sink in water to evaluate their volume by water displacement. Samples were then dried for 48 hours at 80 degrees celsius and weighted.

@@

We calculated five functionnal traits of leaf and wood samples: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the the wood sample dry weight in $g$ divided by wood sample volume in $cm^3$ (see table 1).

Table 1: Functional traits measured. *Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development from @baraloto_decoupled_2010.*

@@

To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river. Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008]. Distance to the nearest potential river was using ArcGis [@desktop_arcgis_2011].

In order to avoid multi-colinearity in future analyses, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement. We reduced our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see supplementary materials S1).

To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density. Canopy height is calculated from a digital canopy model built with a LIDAR campaign on Uppangala **date ?** plots with a resolution $1*1 m^2$. Both basal area and stem density are calculated from 2013 girth inventory of Uppangala plot on stemw with a diameter above 10 centimeters at breast height ($DBH>10$). Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$). So we removed stem density to avoid multicolinearity, and we kept canopy height that was not correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).

@@

*Title are just here to help remembering the structure but might be removed*

### Functionnal composition of communities regarding environment

#### Covariation in traits

Traits variation was first investigated in three levels with the coefficient of variation ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2). Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).

@@

#### Community analysis

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. To investigate traits distribution inside communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010]. Analyses were conducted on the five functionnal traits and the two economic spectra defined with the species scores obtained on the 2 first PCA axes.

CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution. We tested links between CWMs and environment with a first null model randomizing traits among species (permuting raw in `Trait*Species` matrix). **Precise what this model imply**. Linear regressions between CWMs and biotic and abiotic factors after step measure were compared to corresponding 999 null linear regressions from our null models with F-statistic.

CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
We tested links between CWVs and environment with a second null model randomizing abundances among species inside each community but keeping the observed list of species and trait values inside communities. **Precise what this model imply**. The CWV standardized effect size (SES) was then computed comparing each community to their corresponding null communities with the formula :
$$SES = \frac{(CWV_{obs} - CWV_{null})}{\sigma _{null}}$$
where $CWV_{obs}$ is the observed CWV and $CWV_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distribution CWVs. Finally proportion of of CWVs differing from null expectaions was tested with a double tails test; and links between SES of CWVs and environment was investigated by linear regression.

In parrallel to CWVs, eveness of communities was investigated with CVNND with the same null model randomizing abundances among species inside each community but keeping the observed list of species and trait values inside communities. SES were also calculated for CVNNDs. Finally proportion of of CVNNDs differing from null expectaions was tested with a double tails test; and links between SES of CVNNDs and environment was investigated by linear regression.

#### Phylogenetic signal

Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
Phylogenetic signal was globaly assessed for each functionnal trait among the phylogenetic tree by Moran's I statistic.
Phylogenetic signal structure was assessed by a tree tip distances correlogram.

### Spatial structure of community compositions

$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions. Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
In parrallel spatial structure of community functionnal composition was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007. Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix).

## Functionnal composition of communities regarding environment

### Traits and environment

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
L[is.na(L)] <- 0
sel <- intersect(row.names(Q), names(L))
Q <- Q[sel,]
L <- L[sel]
R <- R[row.names(L),]
L <- list(abundance = L, `presence-absence` = L)
L$`presence-absence`[L$`presence-absence` > 1]  <- 1
rm(Deriv, XY, com, Env, Trees, sel, PFT_sp)
```

### Community analysis

@@

# Simulated CWMs
n = 999
nQ <- rep(list(Q), n)
nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
cl <- makeCluster(3)
clusterExport(cl, list('L', 'nQ'))
nCWM <- clusterApply(cl, L, function(l){
  lapply(nQ, function(q){
    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){
      apply(l, 1, function(x){
        weighted.mean(q[,v], x, na.rm=T)
      })
    })
  })
})
stopCluster(cl)
rm(nQ, cl)
nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
## Regressions
nCWM.lm <- mapply(function(x,y){
  lapply(y, function(z){
    mapply(function(r,s){
      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
    }, r = x, s = z, SIMPLIFY = F)
  })
}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
rm(nCWM) ; invisible(gc())
## Global F
nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
nFval <- lapply(nFval, function(f){lapply(f, unlist)})
pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
row.names(pval) <- row.names(Fval)
rm(Fval, nFval)
# Partial F
npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
ppval$abundance$BA <- NA
ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
ppval <- lapply(ppval, function(x){1 - x})
rm(pFval, npFval)

@@

### Phylogenetic signal

@@

Table 7: Summary of Moran’s I statistic and p-values for phylogenetic signal of functionnal traits. _See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

## Spatial structure of community compositions

Spatial structure of both community phylogenetic and functionnal communities don't show significant organization, no convergence nor divergence ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).

**Following tables will probably be removed in final report by simply including results in text. They are juste here for discussion.**

Table 8: Summary of spatial structure of community functionnal compositions. _$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions. Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

Table 9: Summary of spatial structure of community phylogenetic compositions. _$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions. Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix). The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

@@

In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling. Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits. Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*). Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).

Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary. Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$).

@@

## S3: Phylogenetic signal among functionnal traits by correlogram

>>>>

Keywords: functional trait, tropical evergreen forest, Western Ghats, India...

@@

**Study scheme:**

1. Do variations and covariations in functionnal traits reveal species ecological strategies ?
    i. Do functionnal traits vary among species ?
    ii. Do covariations in functional traits represent axes of ecological strategy ?
2. Do functional composition reflect the influence of abiotic and biotic drivers of community dynamics ?
    i. locally
    ii. between communities

@@

* Functional traits can be linked to trade-offs in tree strategy for functionning, survival and reproduction [@baraloto_decoupled_2010, and see table 1], which vary depending on environment condition. For instance, SLA varies along a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
* Leaf investment definition
* Leaf ressource acquisition definition

@@

The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996. The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998]. Each tree girth is monitored by a two-year census.
**Girth monitoring to develop shortly**

@@

We collected trait data on 91 over 117 (78%) species. We sampled five individuals per species for most abundant species and we completed with less abundant and rare species with one to five individuals. Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals. Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot. For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013. Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy. Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling. Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours. For some individuals we were unable to collect an appropriate branch sample (58 over 267 individuals representing 12 missing species over 91).

@@

Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite]. Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time. Once dried, the leaves were weighted directly after been extracted from the oven to avoid humidity bias. The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues. Branch wood samples were sink in water to evaluate their volume by water displacement. Samples were then dried for 48 hours at 80 degrees celsius and weighted.

We calculated five functional traits of leaf and wood samples: leaf thickness (Thick in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).

Table 1: Functional traits measured. *Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*

@@

We selected 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river. Digital elevation model is issued from a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008]. Distance to the nearest potential river was computed using ArcGis [@desktop_arcgis_2011].

In order to avoid multi-colinearity in future analyses, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning. We reduced our explanatory variables to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness to represent monsoon winds [@bunyan_effect_2015] and slope.

We selected two uncorrelated biotic factors: canopy height and basal area (Spearman's $rho = -0.09$, $p-value > 0.1$). Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$. Basal area was calculated from 2013 girth inventory of Uppangala plot on stems with a diameter above 10 centimeters at breast height ($DBH>10~cm$). 
<!-- Use of canopy height and basal area as biotic factors of our 5 functional traits was justified by structural equation modelling (see supplementary material S2). -->

@@

### Traits variation and covariation

Traits variation was first investigated in three levels with the coefficient of variation ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species (inter-specific).

### Functional composition of communities

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. To test for pattern in functional composition of communities we used null model approach. A first null model (NM1) randomizes traits among species (permuting raw in `Trait*Species` matrix). The second model (NM2) randomizes abundances among species inside each community but keeps the observed list of species and trait values inside communities. NM1 allow to assess environmental filtering and biotic process on Uppangala plot species. NM2 bring more insight on environmental and biotic effects on species distribution inside communities [@bernard-verdier_community_2012].

To investigate traits distribution of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010]. CWV lower than expected under a null model will be evidence for convergence of traits inside the community whereas higher CWV will show divergence in trait distribution. CVNND is used to detect niche differenciation patterns in case of a lower CVNND as observed under a null model.

CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution. CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
CVNND was computed in abundance inside each community. For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}}$$
where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.

Linear regressions between CWMs and biotic and abiotic factors after step measure were compared to corresponding 999 null linear regressions from NM1 with F-statistic. For both CWV and CVNND, proportion of of communities differing from null expectaions was tested with a two-tailed test; and links between SES and environment was investigated by linear regression. Analyses were conducted on the five functional traits and the two economic spectra defined with the species scores obtained on the 2 first PCA axes.

### Composition variation between communities

Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1. Phylogenetic signal was globaly assessed for each functional trait among the phylogenetic tree by Moran's I statistic. Positive Moran's I value will indicate convergence of trait value inside the phylogeny whereas negative Moran's I will show divergence.

$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities. Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion. In parrallel variation of functional composition between communities was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007. Observed values were compared to NM1. Negative Hardy's metrics significate that species are functionaly or phylogenetically

## Traits variation and covariation

Trait coefficients of variation analysis reveals a general increase of functional trait variation among scales (from intra-individual to inter-specific, see table 2). Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).

@@

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
L[is.na(L)] <- 0
sel <- intersect(row.names(Q), names(L))
Q <- Q[sel,]
L <- L[sel]
R <- R[row.names(L),]
L <- list(abundance = L, `presence-absence` = L)
L$`presence-absence`[L$`presence-absence` > 1]  <- 1
# rm(Deriv, XY, com, Env, Trees, sel, PFT_sp)
```

## Functional composition of communities

@@

# Simulated CWMs
n = 999
nQ <- rep(list(Q), n)
nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
cl <- makeCluster(3)
clusterExport(cl, list('L', 'nQ'))
nCWM <- clusterApply(cl, L, function(l){
  lapply(nQ, function(q){
    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){
      apply(l, 1, function(x){
        weighted.mean(q[,v], x, na.rm=T)
      })
    })
  })
})
stopCluster(cl)
rm(cl)
nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
## Regressions
nCWM.lm <- mapply(function(x,y){
  lapply(y, function(z){
    mapply(function(r,s){
      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
    }, r = x, s = z, SIMPLIFY = F)
  })
}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
rm(nCWM) ; invisible(gc())
## Global F
nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
nFval <- lapply(nFval, function(f){lapply(f, unlist)})
pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
row.names(pval) <- row.names(Fval)
rm(Fval, nFval)
# Partial F
npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
ppval$abundance$BA <- NA
ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
ppval <- lapply(ppval, function(x){1 - x})
rm(pFval, npFval)

@@

## Composition variation between communities

@@

Table 7: Summary of Moran’s I statistic and p-values for phylogenetic signal of functional traits. _See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

Variations of both functional and phylogenetic compositions between communities don't significantly differ from null models ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).

**Following tables will probably be removed in final report by simply including results in text. They are just here for discussion.**

Table 8: Summary of variation of functional composition between communities. _$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities. Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

Table 9: Summary of variation of phylogenetic composition between communities. _$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities. Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix). The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

## Strategy axes
## Community assembly
## Process scales
## Limits and perspectives

@@

In order to see if or if environmental biotic factors were the results of functional traits we used structural equation modelling. Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *functional traits* defined with our 5 functional traits. Two models were compared, a first one with the functional traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functional traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*). Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).

Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functional traits and not the contrary. Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$).

@@

## S3: Phylogenetic signal among functional traits by correlogram
